The CMS Data Aggregation System

### 2010-05-04 doc/paper.tex
More tune-ups
git-svn-id: svn+ssh://svn.cern.ch/reps/CMSDMWM/DAS/trunk@650 4525493e-7705-40b1-a816-d608a930855b

\begin{abstract}
Meta-data plays significant role in large modern enterprises, 
research experiments and digital libraries where it comes from a different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information can be accessed via
pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine simplicity of keyword-based search with precision of RDMS
under the new system. The aggregated information being collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Such close to real-time system 
based on studies  of use cases of the CMS particle physics experiment at 
the LHC where we used a large scale, distributed computing system 
to motivate this work.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Meta-data\sep data aggregation\sep HEP.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works, see \cite{DBXplorer, QueryAnswer} for more references,
are based on some kind of the graph over the underlying database and
use a generated database out of pristine one to perform the work.
The proximity of the results in those techniques 
are mostly confined to the text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL syntax.

To address those limitations an attempt to build a simple, intuitive and 
flexible query language was introduced in \cite{DBS-QL}.
It represents a power of SQL while hiding underlying relational schema from 
the end-users. As a results a human questions were intuitively mapped into 
simple queries. For example, the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
This solution represents a common use cases within HEP community and provide
intuitive mapping between mental model of end-users and underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be user to 
search desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated DB it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed heterogeneous environment where different security policies
are in place.

@@

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accomodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to 7 Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The 50+ Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services being designed and developed to
maintain detector and production operations, including detector
conditions databases, data location and bookkeeping services,
data transfer and job monitoring tasks. Even though majority of them
are located at CERN, it was never been a requirement in CMS computing
and data model. 
%For instance, the production teams operated at Tier-1,2
%centers set up local services for data bookkeeping and operational
%tasks. Based on their resources the choice of database back-end was
%up to the site managers. Therefore the CMS software were designed to support different
%DB back-ends and provide tools for migrating data across them.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose design, implementation and
security policies are left up to their developers. Therefore DAS does not require
a data preservation and transaction capabilities. Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. Teh DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provides
ability to place queries, similar to SQL, without specific knowledge of 
underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support of keyword based search queries with ability to use condition operators;
\item pluggable interface to existing and up-coming data-service APIs;
\item support heterogeneous software environment and distributed nature of the data-services;
\item preserve security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementations and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query where placed into analytics DB and mapped into set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregated results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
Analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicates with different data-services and retrieves
data on demand. This was done by invoking appropriate data-service plugin
upon the user request. Each plugin consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service DAS map contains information about data-service APIs,
input and output parameters and their mapping to DAS keys. 
This information was stored into DAS Mapping DB which we discuss next.
Below you can see a typical DAS Mapping DB record for CMS
SiteDB data-service \cite{SiteDB}:
%\footnote{We used regular expression 
%patterns, evaluated during run time, to validate provided set of input 
%paremters and DAS keys.}:
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
Although the majority of data-services looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required. For instance,
to deal with data-services which were only accessible via private network.
Therefore, a way to propagate user credentials was required, leading
to specific plugin implementation.

@@

\subsection{DAS Mapping DB}
DAS Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values as regular expression patterns into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used \verb+logical_file_name+ notation to identify the name of the
logical file, while the same entity was called \verb+lfn+ 
in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key \verb+file+.
%\footnote{We
%also deal with physical file names, but still used {\it file} to refer to them in
%DAS notations, since simple regular expressions were used to identify the
%meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type \verb+file+ in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. In aforementioned example, 
usage of \verb+file+ in DAS query cause a call of DBS 
\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
This approach gave power to end-users to express the meaning of their queries, especially
when they deal with numbers, and achieve accuracy of results. 
For instance, users were able to type \verb+run > 10 and run < 100+
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full keyword search is under development.
% within MongoDB.

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions ($conds$) 
and selection keys ($skeys$)
\item Check cache for query hash, if found: look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ identifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previos step
\begin{verbatim}
if cache_has(row.primary_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
We did use a bulk insert operations to increase throughput. The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
\item Get results from the cache
\end{enumerate}
%The data was stored in a form of DAS records into the cache, while
%queries were recorded into Analytics DB for further analysis.
We also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

@@

\subsection{DAS queries}
Apart from previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+. Those
names were used by our users in their daily work operations. So, by typing a
\verb+file+ we didn't perform a string matching to the 
``file'' string, but rather look-up all file objects in a DAS. 
Moreover each DAS key has a default attribute.
For intance, the default attribute for
\verb+dataset+, \verb+block+, \verb+file+ 
was a \verb+name+, while the default attribute for
\verb+run+ and \verb+event+ was a \verb+number+. 
This allows to specify key-value pair as simple as possible, e.g.
\verb+run=123456+, which means run number 123456. At the same time a query 
key, e.g. \verb+site+ meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as \verb+run=123456+ instructs system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.

>>>>

\begin{abstract}
Meta-data plays significant role in large modern enterprises, 
research experiments and digital libraries where it comes from a different 
sources and distributed in a variety of forms and digital formats. 
It is organized and managed by constantly evolving software using 
both relational and non-relational data sources. Even though we can apply
information retrieval approach to non-relation data sources,
we can't do so for relational ones, where information is accessed via
pre-established set of data-services.

Here we discuss a new data aggregation system which consumes, 
indexes and delivers information from different relational and 
non-relational data sources to answer cross data-service queries 
and explore meta-data associated with petabytes of experimental data. 
We combine simplicity of keyword-based search with precision of RDMS
under the new system. The aggregated information being collected from various sources,
allowing end-users to place dynamic queries, get precise answers and 
trigger information retrieval on demand. Such close to real-time system 
based on studies  of use cases of the CMS particle physics experiment at 
the LHC where we used a large scale, distributed computing environment 
to motivate this work.
\end{abstract}

\begin{keyword}
%% keywords here, in the form: keyword \sep keyword
Meta-data\sep data aggregation\sep information discovery\sep HEP.

@@

\section{Related Work\label{RelatedWork}}
Even though the idea of keyword queries over the relational database
%querying relation databases via keyword based search
%algorithms 
is not knew it is still under discussion in computer science domain.
A few alternative solutions has been proposed to address this issue
in last several years. In \cite{DBXplorer} the conjunctive keyword queries,
i.e. retrieval of documents that contain all query keywords, has
been discussed. It requires that all keywords are matched in a row from a
single table or joined tables. Authors of \cite{QueryAnswer} made a step forward
by considering information around matched values and presents customized answers to
end-users over the generated new database. These and other related 
works, see \cite{DBXplorer, QueryAnswer} for more references,
are based on some kind of the graph over the underlying database and
use a generated database out of pristine one to perform the work.
The proximity of the results in those techniques 
are mostly confined to a text-based values stored in a database, since 
usage of numerical values in the input search field is mostly useless without 
knowledge of the context of the value. For example, typing a value 100 
in the search input field leads to plenty of inappropriate matches, 
such as row ids, and in order to be
answered correctly requires additional keyword, clarifying its meaning, 
which should be matched with particular
entity (table.column) in the underlying database.
Therefore to answer presicely the logical questions, such as  
{\it find me total number of files whose size more then 20 but less then 
100} would be hard or almost impossible in \cite{DBXplorer, QueryAnswer}, 
while it can be easily accomplished using SQL syntax.

To address those limitations an attempt to build a simple, intuitive and 
flexible query language was introduced in \cite{DBS-QL}.
It represents a power of SQL while hiding underlying relational schema from 
the end-users. As a results a human questions were intuitively mapped into 
simple queries. For instance, the question
{\it I'm looking for files who contain data taken on certain date and located at
particular site} was represented as simple as \cite{DBS-QL}
\begin{verbatim}
find file where date 2009-01-02 11:59 CET and site = T2
\end{verbatim}
This solution represents a common use cases within HEP community and provide
intuitive mapping between mental model of end-users and underlying database schema.

At the same time the data aggregation is commonly addressed in the context of
federated databases, see for example \cite{FedDB}. It unifies data coming 
from different RDMS systems into the federated DB where SQL queries can be used to 
search desired data. Even though it would be interesting to apply technique 
discussed in \cite{DBS-QL} to a federated DB it still requires 
an access to the underlying database schema. This obviously creates a problem
in distributed, heterogeneous environment where different security policies
are in place.

@@

The CMS distributed computing and data model \cite{CMSDataModel} 
is designed to process
and efficiently manage a few PBs of data expected each year
in CMS operation at LHC. The computing resources provided by members of CMS
collaboration are geographically distributed, 
interconnected via high throughput networks and operated by means 
of Grid software. The model allows to cover broad variety of
hardware, mass and storage elements and configuration of the
clusters. To accomodate efficient data processing CMS uses
a multi-Tier distributed model, where specific tasks of data taking,
processing, archival and distribution are assigned to each tier based
on CMS data model. For example, the Tier-0 center at CERN is responsible
for archiving the data coming out from the detector, prompt first pass reconstruction
and data distribution to Tier-1 centers, located around the world.
%The 7 Tier-1 centers
%located in France, Germany, Ialy, Spain, Taiwan, United Kingdom and United States
%keep a portion (copy) of the data delivered by Tier-0 for further processing.
They provide storage and CPU power for high priority analysis jobs.
The Tier-2 centers
%, located at more then 50 sites around the world,
are dedicated for user analysis tasks and production of simulated data.

A broad variety of data-services have being designed and developed to
maintain detector and production operations, including detector
conditions databases, data location and bookkeeping services,
data transfer and job monitoring tasks. Even though majority of them
are located at CERN, it was never been a requirement in CMS computing
and data model. 
%For instance, the production teams operated at Tier-1,2
%centers set up local services for data bookkeeping and operational
%tasks. Based on their resources the choice of database back-end was
%up to the site managers. Therefore the CMS software were designed to support different
%DB back-ends and provide tools for migrating data across them.

@@

\section{Data Aggregation System\label{DAS}}
The goal of the Data Aggregation System was to combine and
aggregate data from existing data-services
under one umrella, utilizing their access rules, APIs, security policies, etc., and
allowing users to perform queries against them. In contrast, the approaches
discussed in \cite{DBXplorer, QueryAnswer, FedDB} are
enforced a publication step of data into a dedicated database 
or central repositories \cite{iRODS}. Therefore, quite often it is required
that data should be converted into uniform data formats in order enable
aggregation, see \cite{OpenArchive}. In DAS architecture the data are provided
by existing data-services and infrastructure, whose implementation,
data representation and security policies are left intact. Consequently,
DAS does not require a data preservation and transaction capabilities. 
Instead DAS can be treated
as proxy or caching layer where information retrieval are triggered dynamically
by user queries. Such infrastructure allows data sharing across different queries and
data aggregation using appropriate mapping. Below we discuss in details the
DAS architecture.

\subsection{DAS architecture}
The design of the DAS system was based on previous studies of the CMS Data 
Bookkeeping System (DBS) \cite{DBS, DBS07}. The DBS Data Discovery service
\cite{DD} was based on DBS Query Language (DBS-QL) \cite{DBS-QL}. It provided
a syntax similar to SQL, without specific knowledge of underlying DB schema.
%We carefully analyzed user
%requests and queries placed into the DBS, their patterns, frequencies and latency. 
%As a result we created a data discovery service \cite{DD}. Originally it used MVC
%architecture and later, being re-factored as pure presentation layer on top of 
%the DBS Query Language (DBS-QL) \cite{DBS-QL} with support of dynamic queries against
%DBS back-ends. The DBS-QL \cite{DBS-QL} syntax was based on SQL query language, where 
%we took away relations and details on underlying schema from end-users. 
%This was achieved by using Dijkstras shortest path algorithm 
%to establish necessary table joins based on provided information in user queries and
%knowledge of foreign key relationships from auto-loaded database schema.
%The DBS-QL keys exposed to end-users were choosen based on a jargon used by
%physicists. They're internally mapped into database entitites and attributes, 
%e.g. the {\it run} DBS-QL key refers to run number attribute of run summary table.
A quick adoption and wide usage of DBS-QL in CMS gave us confidence in choosen approach 
and provide a basis for building DAS system.

As was pointed out in \cite{Arms} a mixed content and 
mixed meta-data and meta-data consistency should be considered as a whole in design 
of the system to successful information discovery. 
Starting from this ground we designed DAS system as an
additional layer on top of the existing data-services
within CMS computing infrastructure by imposing the following set of requirements:
\begin{itemize}
\item support a keyword based search queries with ability to use condition operators;
%\item pluggable interface to existing and up-coming data-service APIs;
\item use existing heterogeneous software environment and distributed nature of the data-services;
\item preserve access rules and security policies of individual data-services;
%\item be flexible to adjust to dynamic nature of data-service APIs, 
%due to continuously changing detector and software environments;
\item retrieve data on demand and aggregate them if necessary across
multiple data-services;
\item be transparent to data-services implementation and their data formats;
%\item support legacy data-services, applications and APIs.
\end{itemize}
\begin{figure}[htb]
\centering
\includegraphics[width=100mm]{DAS_Cache_and_Analytics.pdf}
\caption{
The DAS architecture diagram. 
Input query were placed into analytics DB and mapped into a set of
data-service APIs and their parameters. If superset of the query has been
found in DAS cache, the appropriate results were delivered back to the users.
Otherwise a data-service APIs were invoked, triggering DAS cache population. 
Once data become available and aggregated in DAS cache, users were able 
to see the results. This on demand system was supplemented by DAS robots 
(unix daemons) who were consulting DAS analytics database to enable data 
population into DAS cache for most common query requests.
}
\label{DAS_cache}
\end{figure}

@@

The DAS architecture is shown in Fig. \ref{DAS_cache}. It consists of
the core library with support of pluggable modules for data retrieval;
caching layer to store data-service output and aggregate results;
request service to handle user queries;
Mapping DB to keep information about data-service APIs, their
notations and mapping to DAS;
Analytics DB for query analysis and identification of pre-fetch 
strategies.

\subsection{DAS core}
The DAS core communicated with different data-services and retrieved
data on demand. This was done by invoking appropriate data-service plugin(s)
upon the user request. Each plugin consists of DAS map, the
data parser(s) to convert data-service output into DAS records
and it was responsible for data-service API calls.
The data-service plugin contains a map about data-service APIs,
input and output parameters and their mapping to DAS keys. For example
\begin{verbatim}
{"api":{"params":{"name":""}, "name":"CMSNametoSE"}, "system":"sitedb",
 "api2das": [{"pattern": "re.compile('^T[0-3]_')", 
              "api_param": "name", "das_key": "site"}], 
 "daskeys": [{"pattern": "re.compile('^T[0-3]_')", 
              "key": "site", "map": "site.name"}]}
\end{verbatim}
This information was stored into DAS Mapping DB which we discuss next.

Although the majority of data-service plugins looks alike
and their logic was abstracted, including parsers, there were a few
cases where special care was required, e.g. access to data-services 
accessible via private network. Therefore, a way to propagate 
user credentials was required, leading to specific plugin implementation.

@@

\subsection{DAS Mapping DB}
The Mapping DB holded the mapping between data-service APIs and DAS keys exposed to
the users. It created a necessary glue between data-services and identifies
how their data can be related to each other. Basically Mapping DB was positioned 
as a {\it registration} service with ability to translate notations among data-services.
Upon identification of APIs which will participate in DAS activity 
we stored their names, input parameters, together with their type and accept 
values, as regular expression patterns, into the Mapping DB. 
Also we stored the mapping between data-service notations and DAS keys. 
For instance, the CMS DBS system \cite{DBS}
used a \verb+logical_file_name+ notation to identify the name of the
logical file, while the same entity was called \verb+lfn+ 
in CMS PhEDEx system \cite{PhEDEx}.
Since both of them were referred to the logical file name, 
we map both notations into DAS key \verb+file+.
%\footnote{We
%also deal with physical file names, but still used {\it file} to refer to them in
%DAS notations, since simple regular expressions were used to identify the
%meaning of the {\it file} keyword from the provided value.}
Such keys, e.g. \verb+file+, were exposed to the end-users as pre-defined keywords to
be used in search queries. For instance, end-users were able to type \verb+file+ in
order to get all file names known to DAS via data-services as well as
provide a concrete patterns associated with a file, e.g. \verb+file = abc*+.
Under the hood, the appropriate data-service APIs were looked up along with appropriate
parameters to answer those questions. In aforementioned example, 
usage of \verb+file+ in DAS query cause a call of DBS 
\verb+listFiles+ and PhEDEx \verb+fileReplicas+ APIs, respectively.
This approach gave ability to end-users to express the meaning of their queries, especially
when they deal with numbers, and achieve accuracy of results. 
For instance, users were able to type \verb+run > 10 and run < 100+
to specify a range of run numbers to be looked up in DAS cache. At the same time
a full keyword search is under development.
% within MongoDB.

@@

\subsection{DAS caching system}
The DAS caching system was used to dynamically fetch
and aggregate data upon user requests. We used the following 
algorithm:
\begin{enumerate}[1.]
\item Decompose query into set of conditions (\verb+conds+) 
and selection keys (\verb+skeys+)
\item Check cache for query hash, if found, look-up results from the cache
\item Identify list of APIs which can serve requested data
\begin{verbatim}
apilist = [r(url, api, args) for r in mapping_db(skeys, conds)]
\end{verbatim}
Here \verb+mappind_db+ defines translation between input query and data-service
APIs which can delivery such data.
\item Iterate over APIs
\begin{verbatim}
for url, api, args in apilist:
    if  superset_in_analytics(api, args):
        continue
    for row in parser(get_data(url, api, args)):
        yield row
\end{verbatim}
Here \verb+superset_in_analytics+ verifies if provided \verb+api, args+
are subset of already registered in Analytics DB api call(s). The 
\verb+get_data+ and \verb+parser+ are define methods to fetch and parse
data, respectively.
\item Perform aggregation by iterating over records from the previos step
\begin{verbatim}
if cache_has(row.primary_key):
   existing_record = get_from_cache(row.primary_key)
   merge(row, existing_record)
   insert_into_cache(row)
   delete_from_cache(existing_record)
else:
   insert_into_cache(row)
\end{verbatim}
The \verb+primary_key+
is a key which identify the record defined in Mapping DB, e.g. \verb+file.name+
for \verb+file+ record.
We did use a bulk insert operations at this step to increase throughput. 
\item Get results from the cache
\end{enumerate}
%The data was stored in a form of DAS records into the cache, while
%queries were recorded into Analytics DB for further analysis.
We also had an option to run independent set of daemons, DAS robots, who
monitor Analytics DB for most popular requests and pre-fetch the most
common data into DAS cache. 
%The interesting part was being able
%to identify the superset of api/args out of found one in an Analytics DB.

@@

\subsection{DAS queries}
Apart from the previous DBS-QL syntax we decided to drop its syntax and allow users
express their queries in a free text-based form. Although we must clarify its
meaning. We used a pre-defined set of keys registered in DAS to describe common
entities of our data. For example, \verb+dataset+, 
\verb+file+, \verb+block+, \verb+run+, \verb+event+. Those
names were used by our users in their daily work operations. So, by typing a
\verb+file+ we didn't perform a string matching to the 
``file'' string, but rather look-up all file objects in a DAS. 
Moreover each DAS key has a default attribute.
For intance, the default attribute for
\verb+dataset+, \verb+block+, \verb+file+ 
was a \verb+name+, while the default attribute for
\verb+run+ and \verb+event+ was a \verb+number+. 
This allows to specify key-value pair as simple as possible, e.g.
\verb+run=123456+, which means run number 123456. At the same time a query 
key, e.g. \verb+site+ meant to look-up all site objects. Such ability
to use key-value pairs resolve ambiguity in object values. For instance,
specifying a run number as \verb+run=123456+ instructs system to resolve
data object with run equal to 123456 rather then other objects which may have
123456 values, e.g. event number or file size. At the same time, to complement
the stregth of the DAS we do plan to index data objects and add a keyword-based 
search later.
